---
title: 'Assignment 2: Inferential Data Analysis'
author: "Pallavi Gadgil"
date: "7/1/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## Inferential Data Analysis

This report contains Inferential Data Analysis of the ToothGrowth Data in the R datasets package. The objective for this assignment is to 

  >1. Load the ToothGrowth data and perform some basic exploratory data analyses
  >2. Provide a basic summary of the data.
  >3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. (Only use the techniques from class, even if there's other approaches worth considering)
  >4. State your conclusions and the assumptions needed for your conclusions.


## Load the ToothGrowth data and perform some basic exploratory data analyses
The ToothGrowth data is loaded from the datasets package. Per R documentation, the data represent "The Effect of Vitamin C on Tooth Growth in Guinea Pigs".

```{r ToothGrowth}
data(ToothGrowth)
head(ToothGrowth)
```

## Provide Basic Summary of the Data

```{r summary}
str(ToothGrowth)
summary(ToothGrowth)
```

As seen in the suumary of this dataset, it has 60 observations on 3 variables.

  >* [,1]	len	numeric	Tooth length
  >* [,2]	supp	factor	Supplement type (VC or OJ).
  >* [,3]	dose	numeric	Dose in milligrams/day

To look at this data further, a box plot is prepared to look at the Tooth growth by Supplement Type and the dosage.
```{r}
qplot(supp, len, data = ToothGrowth, 
      facets = ~dose, 
      main = "Tooth growth by supplement type and dosage", 
      xlab = "Supplement type", 
      ylab = "Tooth Growth Length") + 
  geom_boxplot(aes(fill = supp))
```


Based on the plot, it looks like the doage may have an impact on the Tooth growth but not quite clear if summplement type has any impact.


##  Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

Lets start with the null hypothesis that supplement type does not have any impact on the tooth growth rate.
```{r suppHypothesis}
t.test(len~supp, paired=FALSE,data=ToothGrowth,var.equal=TRUE)
```
As we can see, the P value is > 0.05 and the 95% confidence interval includes 0. Hence there is not enough evidence to reject the null Hypothesis.

### Lets start comparing growth rate for dosage. 

Since the dosage has 3 unique values (0.5,1.0 and 2.0), we will split the data set in 2 dosage types to run the t tests. Each comparision set will have null hpothesis that the increase in the dosage does not increase the tooth growth length. 


#### Comparing dosage 0.5 and 1.0

```{r dose510}
toothGrowth0510<- subset(ToothGrowth, dose %in% c(0.5,1.0))
t.test(len~dose, paired=FALSE,data=toothGrowth0510,var.equal=TRUE)
```

#### Comparing dosage 0.5 and 2.0
```{r dose520}
toothGrowth0520<- subset(ToothGrowth, dose %in% c(0.5,2.0))
t.test(len~dose, paired=FALSE,data=toothGrowth0520,var.equal=TRUE)

```

#### Comparing dosage 1.0 and 2.0
```{r dose1020}
toothGrowth1020<- subset(ToothGrowth, dose %in% c(1.0,2.0))
t.test(len~dose, paired=FALSE,data=toothGrowth1020,var.equal=TRUE)
```

As we can see, all 3 tests have close to 0 p value and the confidence interval at 95% does not cross over 0. Hence we must reject the null hypothesis and accept the alternative,i.e.  Tooth Growth increases with an increasing dose.

## Conclusion
Based on the above t tests we can conclude that
>1.  Supplement type has no effect on tooth growth/length
>2. Tooth Growth increases with an increasing dose.


